Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-15 doc/manuscript/ESSDmanuscript.Rmd
update variable mapping
added woody.mortality and woody.mortality_root. Still need to add these to ForC_variables.

```{r fig_variable_mapping, out.width = "15cm", echo = FALSE, fig.cap = "\\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., minimum DBH). Additional caveats are as follows: 1- assumes that change in foliage biomass is negligible (see note 7); 2- incomplete: excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- incomplete: excludes belowground components;  4-incomplete: excludes breakage into pieces less than dead wood threshold size; 5-incomplete: excludes woody mortality of stems <10 cm DBH, decomposition of dead wood (aboveground and coarse roots) into sizes classified as litter, may exclude branch fall; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change."}
#this is for a two-column figure. for one-column figure, use out.width = "8.3cm"
knitr::include_graphics("figures_tables/C_variable_mapping.png")
```

@@

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table, format = "latex",
             longtable = TRUE,
             caption = "\\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. ") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)  %>%
  row_spec(1, bold = TRUE)  %>% #biomass
  row_spec(5, bold = TRUE)  %>% #agb
  row_spec(13, bold = TRUE) %>% #root
  row_spec(21, bold = TRUE)  %>% #dead wood
  row_spec(28, bold = TRUE)  %>% #litter
  row_spec(35, bold = TRUE)  %>% #som
  row_spec(39, bold = TRUE) #total
```

>>>>

```{r fig_variable_mapping, out.width = "15cm", echo = FALSE, fig.cap = "\\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., min DBH). Additional caveats are as follows: 1- often excludes branch fall, assumes that change in foliage biomass is negligible (see note 7); 2- excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- excludes belowground components;  4-excludes breakage into pieces less than dead wood threshold size; 5- may exclude woody mortality of stems <10 cm DBH and branch fall, decomposition of dead wood into sizes classified as litter; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change."}
#this is for a two-column figure. for one-column figure, use out.width = "8.3cm"
knitr::include_graphics("figures_tables/C_variable_mapping.png")
```

@@

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table, format = "latex",
             longtable = TRUE,
             caption = "\\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. ") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)  %>%
  row_spec(1, bold = TRUE)  %>% #biomass
  row_spec(6, bold = TRUE)  %>% #agb
  row_spec(14, bold = TRUE) %>% #root
  row_spec(25, bold = TRUE)  %>% #dead wood
  row_spec(33, bold = TRUE)  %>% #litter
  row_spec(40, bold = TRUE)  %>% #som
  row_spec(44, bold = TRUE) #total
```

### 2022-04-15 doc/manuscript/ESSDmanuscript.tex
update variable mapping
added woody.mortality and woody.mortality_root. Still need to add these to ForC_variables.

\begin{figure}
\includegraphics[width=15cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., minimum DBH). Additional caveats are as follows: 1- assumes that change in foliage biomass is negligible (see note 7); 2- incomplete: excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- incomplete: excludes belowground components;  4-incomplete: excludes breakage into pieces less than dead wood threshold size; 5-incomplete: excludes woody mortality of stems <10 cm DBH, decomposition of dead wood (aboveground and coarse roots) into sizes classified as litter, may exclude branch fall; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change.}\label{fig:fig_variable_mapping}
\end{figure}

@@

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. }\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass &  &  &  & \\
\hline
delta.biomass &  &  &  & \\
\hline
NPP\_woody &  &  &  & \\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag &  &  &  & \\
\hline
biomass\_ag\_woody &  &  &  & \\
\hline
biomass\_ag\_foliage &  &  &  & \\
\hline
delta.agb &  &  &  & \\
\hline
ANPP\_woody\_stem &  &  &  & \\
\hline
ANPP\_foliage &  &  &  & \\
\hline
woody\_mortality &  &  &  & \\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root &  &  &  & \\
\hline
biomass\_root\_fine &  &  &  & \\
\hline
biomass\_root\_coarse &  &  &  & \\
\hline
delta.biomass\_root &  &  &  & \\
\hline
BNPP\_root &  &  &  & \\
\hline
BNPP\_root\_fine &  &  &  & \\
\hline
BNPP\_root\_coarse &  &  &  & \\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood &  &  &  & \\
\hline
deadwood\_standing &  &  &  & \\
\hline
deadwood\_down &  &  &  & \\
\hline
delta.deadwood &  &  &  & \\
\hline
delta.deadwood\_standing &  &  &  & \\
\hline
delta.deadwood\_down &  &  &  & \\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon &  &  &  & \\
\hline
delta.O.horizon &  &  &  & \\
\hline
litter &  &  &  & \\
\hline
delta.litter &  &  &  & \\
\hline
NPP\_litterfall &  &  &  & \\
\hline
R\_het\_litter &  &  &  & \\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC &  &  &  & \\
\hline
delta.SOM / delta.SOC &  &  &  & \\
\hline
R\_het\_soil &  &  &  & \\
\hline
\textbf{TOTAL} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
\end{longtabu}

>>>>

\begin{figure}
\includegraphics[width=15cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., min DBH). Additional caveats are as follows: 1- often excludes branch fall, assumes that change in foliage biomass is negligible (see note 7); 2- excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- excludes belowground components;  4-excludes breakage into pieces less than dead wood threshold size; 5- may exclude woody mortality of stems <10 cm DBH and branch fall, decomposition of dead wood into sizes classified as litter; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change.}\label{fig:fig_variable_mapping}
\end{figure}

@@

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. }\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass &  &  &  & \\
\hline
delta.biomass &  &  &  & \\
\hline
NPP\_woody &  &  &  & \\
\hline
woody.mortality &  &  &  & \\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag &  &  &  & \\
\hline
biomass\_ag\_woody &  &  &  & \\
\hline
biomass\_ag\_foliage &  &  &  & \\
\hline
delta.agb &  &  &  & \\
\hline
ANPP\_woody &  &  &  & \\
\hline
ANPP\_foliage &  &  &  & \\
\hline
woody.mortality\_ag &  &  &  & \\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root &  &  &  & \\
\hline
biomass\_root\_fine &  &  &  & \\
\hline
biomass\_root\_coarse &  &  &  & \\
\hline
delta.biomass\_root &  &  &  & \\
\hline
delta.biomass\_root\_coarse &  &  &  & \\
\hline
delta.biomass\_root\_fine &  &  &  & \\
\hline
woody.mortality\_root &  &  &  & \\
\hline
BNPP\_root\_fine &  &  &  & \\
\hline
BNPP\_root.turnover\_fine &  &  &  & \\
\hline
BNPP\_root\_coarse &  &  &  & \\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood &  &  &  & \\
\hline
deadwood\_standing &  &  &  & \\
\hline
deadwood\_down &  &  &  & \\
\hline
delta.deadwood &  &  &  & \\
\hline
delta.deadwood\_standing &  &  &  & \\
\hline
delta.deadwood\_down &  &  &  & \\
\hline
R\_het\_deadwood & 0 &  &  & \\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon &  &  &  & \\
\hline
delta.O.horizon &  &  &  & \\
\hline
litter &  &  &  & \\
\hline
delta.litter &  &  &  & \\
\hline
NPP\_litterfall &  &  &  & \\
\hline
R\_het\_litter &  &  &  & \\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC &  &  &  & \\
\hline
delta.SOM / delta.SOC &  &  &  & \\
\hline
R\_het\_soil &  &  &  & \\
\hline
\textbf{TOTAL} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
\end{longtabu}
